Holomorphic field theories and higher algebra

### 2024-09-26 draft.tex
update

Let $X$ be an algebraic variety defined over $\CC$, equipped with a symplectic form $\omega_X$.
Then we posit for the relevant phase space, the ``symplectic space'' $LX$---the formal loop space of $X$.
This is the space of maps $\Hat{D}^\times \to X$ from the formal punctured disk to $X$.
%
%Let $V$ be a finite-dimensional complex vector space, equipped with a symplectic pairing $\omega$ that is complex-valued.
%We can view $V$ as $T^* L$ for some Lagrangian vector subspace $L \subset V$, if we wish.
%Then we posit for the relevant phase space, the ``symplectic complex manifold'' $V[z,z^{-1}]$,
%by which we mean the $V$-valued Laurent polynomials in a coordinate~$z$.
%The space encodes algebraic maps from the variety $\CC^\times = \CC \setminus 0$ to $V$,
%so we view it as a kind of algebraic loop space of~$V$.
(There are clearly variations on this idea, such as the holomorphic loop space of all holomorphic maps from $\CC^\times$ to~$X$, when $X$ is a complex manifold.)

@@

The subspace of ``contractible formal loops" $V[[z]]$ inside $V((z))$ produces a module for $\cO(V((z)))$ given by $\cO(V[[z]])$. 
(This module is in fact a quotient algebra.)
One can quantize this module to a module ${\rm Vac}[V]$ for $W[V]$,
where as a vector space, it is still $\cO(V[[z]])$.
This module structure means there is a map of algebras $W[V] \to \End({\rm Vac}[V])$,
so we have a natural ``Hilbert space'' or Fock module for~$W[V]$.

@@

@@

The story we have just told admits a natural extension to higher complex dimensions,
but it requires a first step in the direction of derived geometry.

@@

\owen{We discussed having the following:
\begin{itemize}
\item Quick overview for topological field theories (=analytic content is de Rham complex): 
\begin{itemize}
\item Axelrod-Singer-Kontsevich method via compactified configuration spaces (Fulton-MacPherson) guarantees no divergences for individual graphs and organizes/clarifies analysis of anomalies
\item Explain what this means in terms of Theorem 3.6 (Elliott-Safronov) and Corollary 4.6 (about existence of functorial field theories)
\item Mention the connection with formality of the $E_d$ operads and related outgrowths in topology 
\end{itemize} 
\item Overview the history of the holomorphic case, with citations
\item Describe a version of Minghao's results (and those of Gaiotto {\it et al}) that emphasizes no divergences for individual graphs and the consequences for anomalies (e.g., relevant wheels for 1-loop and higher loop generalization via Laman graphs)
\item Explain what this means in terms of Theorem 3.9 (existence of polydiscs algebras)
\end{itemize}}

\owen{Old copy that might be recycled into final version}

So far we have not discussed the issue of quantization and renormalization for holomorphic field theories.
In the topological setting, renormalization is very well-controlled.
The well-combed configuration space method pioneered by \cite{} has allowed for a rigorous construction of the path integral for many topological field theories.

In other words, by a judicious choice of gauge-fixing, 
we find that the one-loop Feynman diagrams have no UV divergences.
Moreover, no higher loop diagrams appear (for combinatorial reasons),
and the action functional satisfies the quantum master equation.
See Section~\ref{sec: susytwist revisited} for some further discussion.

Although holomorphic field theories had already appeared in practice,
my collaborator Brian Williams was the first to lay a systematic, mathematical foundations (cf. his thesis \cite{BWthesis} and its offshoot \cite{BWhol}).
He did the following:
\begin{itemize}
\item established key analytic results about holomorphic renormalization, 
showing that it is highly manageable, with no counterterms to 1-loop; and
\item characterized the one-loop anomalies  (i.e., obstructions to BV quantization).
\end{itemize}
These apply to all holomorphic theories on $\CC^n$,
but the techniques typically extend to more general complex manifolds.

\brian{Brian's stuff below}

The idea of renormalization is a necessary ingredient in producing examples of typical (Riemannian or Lorentzian) quantum field theories.
One can apply the same sort of techniques to produce holomorphic quantum field theories, and the higher algebraic structures associated to them.
A key feature of holomorphic theories is the absence of (possibly scheme-dependent) counterterms which can make typical quantum field theories difficult to access.

@@

>>>>

Let $X$ be an algebraic variety defined over $\CC$, equipped with a symplectic form $\omega_X$.
Then we posit for the relevant phase space, the ``symplectic space'' $LX$---the formal loop space of~$X$.
This is the space of maps $\Hat{D}^\times \to X$ from the formal punctured disk to $X$.
%
%Let $V$ be a finite-dimensional complex vector space, equipped with a symplectic pairing $\omega$ that is complex-valued.
%We can view $V$ as $T^* L$ for some Lagrangian vector subspace $L \subset V$, if we wish.
%Then we posit for the relevant phase space, the ``symplectic complex manifold'' $V[z,z^{-1}]$,
%by which we mean the $V$-valued Laurent polynomials in a coordinate~$z$.
%The space encodes algebraic maps from the variety $\CC^\times = \CC \setminus 0$ to $V$,
%so we view it as a kind of algebraic loop space of~$V$.
(There are clearly variations on this idea, such as the holomorphic loop space of all holomorphic maps from $\CC^\times$ to~$X$, when $X$ is a complex manifold.)

@@

The subspace of ``contractible formal loops'' $V[[z]]$ inside $V((z))$ produces a module for $\cO(V((z)))$ given by $\cO(V[[z]])$. 
(This module is in fact a quotient algebra.)
One can quantize this module to a module ${\rm Vac}[V]$ for $W[V]$,
where as a vector space, it is still $\cO(V[[z]])$.
This module structure means there is a map of algebras $W[V] \to \End({\rm Vac}[V])$,
so we have a natural ``Hilbert space'' or Fock module for~$W[V]$.

@@

We were driven, so far, by an example which started from the same data as in the non-holomorphic---the data of a symplectic vector space $V$.
For another very natural example, one can consider a vector space $U$ which is equipped with a \textit{symmetric} pairing $\<-,-\>_U$.
Then, the loop space $L(U^*)$ is equipped with a bivector
\beqn
(f,g) \mapsto \underset{z=0}{\text{Res}} \, \<f, \del g\>_U .
\eeqn
Notice that this bivector does \textit{not} correspond to a symplectic form, but nevertheless we can contemplate the deformation quantization of the $\infty$-dimensional Poisson space $L(U^*)$---or, more invariantly, the space of one-forms on the formal disk with values in $U^*$.
The quantization of this Poisson algebra is the algebra with linear generators $f,g \in U((z))$ satisfying the relation
\beqn
[f,g] = \<f, \del g\>_U .
\eeqn
In the case $U=\C$, this is the associative algebra underlying the free chiral boson CFT.

@@

The story we have just told admits a natural extension to higher complex dimensions,
but it requires a first step in the direction of derived geometry.
The main idea is that we want to replace the algebra of Laurent series of a single variable with an appropriate algebra involving more than one variable.

@@

Renormalization is a key ingredient in Riemannian or Lorentzian quantum field theories.
One can apply the same sort of techniques to produce holomorphic quantum field theories, and the higher algebraic structures associated to them.
A key feature of holomorphic theories is the absence of (possibly scheme-dependent) counterterms which can make typical quantum field theories difficult to access.

@@

\subsection{Heat kernel renormalization for holomorphic field theories}
